---
title: "Mizrak et al data -- process replicate 1"
author: "Selin Jessa"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: show
    css: ../../../sjessa/from_hydra/misc/rmd_custom.css
    theme: flatly
    toc: yes
    toc_depth: 4
    number_sections: true
    df_print: paged
---

```{r config, warning = FALSE}

# **MODIFY THIS CHUNK**

project_id <- "mizrak_data" # determines the name of the cache folder
doc_id     <- "01" # determines name of the subfolder of `figures` where pngs/pdfs are saved
out        <- paste0("../processed_data/", doc_id)
figout     <- paste0("figures/", doc_id, "/")
cache      <- paste0("~/tmp/", project_id, "/", doc_id, "/")

```

```{r setup, include = FALSE}

# NO NEED TO MODIFY THIS CHUNK

knitr::knit_hooks$set( echo_label = function(before, options, envir) {
  if ( before ) {
    # Do nothing
  } else sprintf('<br><span style="color:#1abc9c">~[output @ *%s*]~</span>',
                 paste0(options$fig.path, "/", options$label, "...") )
})

knitr::opts_chunk$set(message = TRUE,
                      warning = FALSE,
                      error = FALSE,
                      cache = TRUE,
                      cache.path = cache,
                      fig.path = figout,
                      fig.keep = "all",
                      dev = c("png", "pdf"),
                      cache.lazy = FALSE)

grDevices::pdf.options(useDingbats = FALSE)

options(knitr.table.format = "html") 
knitr::opts_knit$set(width = 1600)

```

***

# Overview

Load expression matrix obtained from GEO, and create a Seurat object in order
to run basic analyses (normalization, dimensionality reduction, clustering, derivation of cluster
markers, etc).

# Libraries

```{r libs, warning = FALSE, message = FALSE, echo = TRUE, cache = FALSE}

library(tidyr)
library(dplyr)
library(readr)
library(magrittr)
library(glue)
library(purrr)
library(ggplot2)
library(cytobox)
library(cowplot)
library(Seurat)

source("../../../sjessa/from_hydra/HGG-G34/samples/functions.R")
source("../../../sjessa/from_hydra/misc/sjlib.R")
ggplot2::theme_set(cytobox::theme_min(border_colour = "black", border_size = 0.5))

```

# Read data

```{r read_data}

id  <- read.table("../data/GSE109447_29319_cells_id_repinfo.txt.gz", stringsAsFactors = FALSE)

# It seems that to get the right cell #s, we leave out c("cell")
label <- read.table("../data/GSE109447_Rep1_29319cells_Basic.txt.gz", stringsAsFactors = FALSE) %>% 
  filter(V1 != "cell")

barcodes <- readxl::read_xlsx("../data/GSE109447_29319_Cell_Barcodes.xlsx", col_names = FALSE)

mat <- data.table::fread("../data/GSE109447_29319_cells.matrix.txt.gz", data.table = FALSE)

mat[1:5, 1:5]

# Separated out the rows containing gene name duplciates
gene_instances <- table(mat[, 2])
duplicates     <- names(gene_instances)[gene_instances > 1]
mat_dups       <- mat[mat[, 2] %in% duplicates, ]
mat_dups$V1    <- NULL

# Sum counts per duplicated gene
mat_dups_summed <- purrr::map(duplicates, function(gene) {
  
  mat_dups[mat_dups$V2 %in% gene, ] %>% 
    select(-V2) %>% 
    colSums()
  
})

mat_dups_summed <- do.call(rbind, mat_dups_summed)
rownames(mat_dups_summed) <- duplicates

# Prep the matrix of non-duplicated counts in the same format
mat_nodups    <- mat[!(mat[, 2] %in% duplicates), ]
mat_nodups$V1 <- NULL
rownames(mat_nodups) <- mat_nodups$V2
mat_nodups$V2 <- NULL
mat_nodups <- as.matrix(mat_nodups)

# Put the two matrices back together
mat_fixed <- rbind(mat_dups_summed, mat_nodups)

# Put it in the same order as the original gene list
mat_fixed <- mat_fixed[unique(mat[, 2]), ]
colnames(mat_fixed) <- paste0(id[[1]], "_", barcodes$...1)

# Create metadata
metadata <- data.frame(ID = id[[1]],
                       Sample = id %>% separate(V1, into = c("rep", "number"), sep = "\\.") %>% pull(rep),
                       Cluster = label[[1]])
rownames(metadata) <- paste0(id[[1]], "_", barcodes$...1)

# Palette
palette <- c("FlatRep1" = "#CBC9E2",
             "FsepRep1" = "#9E9AC8",
             "MsepRep1" = "#BAE4B3",
             "MlatRep1" = "#74C476")

```

# QC

## Create the Seurat object

```{r seurat}

seurat <- CreateSeuratObject(raw.data = mat_fixed,
                             min.cells = 3,
                             min.genes = 0,
                             is.expr = 0,
                             project = "Mizrak_VSVZ",
                             meta.data = metadata)

# Add mitochondrial content to the metadata
mito_genes <- grep("^mt-", rownames(seurat@data), value = TRUE)
percent_mito <- Matrix::colSums(seurat@data[mito_genes, ])/Matrix::colSums(seurat@data) * 100
seurat <- AddMetaData(seurat, percent_mito, "percent_mito")

# Add cell cycle score
seurat <- addCellCycle(seurat, species = "m_musculus")

# Filter doublets
seurat <- SubsetData(seurat, cells.use = rownames(metadata[metadata$Cluster != "Doublet", ]))

```

## QC metrics

```{r qc_before_filt, fig.width = 10, fig.height = 5}

ngene_lim <- 3500
numi_lim  <- 10000
mito_lim  <- 10

plot_grid(VlnPlot(seurat, c("nGene"), cols.use = palette,       
                  do.return = TRUE, group.by = "Sample", point.size.use = -1, x.lab.rot = TRUE) + ylim(NA, ngene_lim),
          VlnPlot(seurat, c("nUMI"),  cols.use = palette,           
                  do.return = TRUE, group.by = "Sample", point.size.use = -1, x.lab.rot = TRUE) + ylim(NA, numi_lim),
          VlnPlot(seurat, c("percent_mito"),cols.use = palette,   
                  do.return = TRUE, group.by = "Sample", point.size.use = -1, x.lab.rot = TRUE) + ylim(NA, mito_lim),
          ncol = 3)

```

# Dimensionality reduction and clustering

```{r process1, results = "hide", dependson = 'thresholds'}

(process_params <- data.frame(
  n_pcs = 10,
  resolution = 0.6,
  seed = 100
))

seurat <- seurat %>%
  NormalizeData(normalization.method = "LogNormalize", scale.factor = 10000) %>%
  FindVariableGenes(mean.function = ExpMean, dispersion.function = LogVMR,
                    x.low.cutoff = 0.0125, x.high.cutoff = 3, y.cutoff = 0.5) %>%
  ScaleData(vars.to.regress = c("nUMI", "percent_mito")) %>%
  RunPCA(pc.genes = .@var.genes, pcs.compute = 100, do.print = FALSE) %>%
  RunTSNE(dims.use = 1:process_params$n_pcs, do.fast = TRUE, seed.use = process_params$seed) %>% 
  FindClusters(dims.use = 1:process_params$n_pcs, resolution = process_params$resolution, random.seed = process_params$seed)

# Set the identity to the provided cluster
seurat <- SetAllIdent(seurat, "Cluster")

seurat <- set_qual_pal(seurat)

saveRDS(seurat, glue("{out}-seurat.Rds"))
write_tsv(process_params, glue("{out}-process_params.tsv"))

```

## Visualization and QC

How many cells per cluster?

```{r n_cells, dependson = 'process1'}

table(seurat@ident)

```

```{r plot_tsne_qc1, fig.width = 11, fig.height = 9, dependson = 'process1'}

plot_grid(cytobox::pca(seurat, title = "PCA", legend = TRUE),
          tsne(seurat, title = "tSNE", legend = TRUE),
          tsne(seurat, title = "nGene", colour_by = "nGene", colours = rdbu, colour_by_type = "continuous", label = FALSE),
          tsne(seurat, title = "nUMI", colour_by = "nUMI", colours = rdbu, colour_by_type = "continuous", label = FALSE),
          tsne(seurat, title = "percent_mito", colour_by = "percent_mito", colours = rdbu, colour_by_type = "continuous", label = FALSE),
          tsne(seurat, title = "G1_S_score", colour_by = "G1_S_score", colours = ylrd, colour_by_type = "continuous", label = FALSE),
          tsne(seurat, title = "G2_M_score", colour_by = "G2_M_score", colours = ylrd, colour_by_type = "continuous", label = FALSE))

```


## PCA

```{r viz_pca, fig.width = 12, fig.height = 4, dependson = 'process1'}

plot_grid(cytobox::pca(seurat, dim1 = 1, dim2 = 2),
          cytobox::pca(seurat, dim1 = 1, dim2 = 3),
          cytobox::pca(seurat, dim1 = 1, dim2 = 4), ncol = 3)

VizPCA(seurat, 1:4, nCol = 4, font.size = 0.9)

getVarianceExplained(seurat)

```

# Markers

Call markers based on the provided cluster labels:

```{r seurat_markers, dependson = 'process1', eval = TRUE}

seurat_markers <- FindAllMarkers(seurat)

write_tsv(seurat_markers, glue("{out}-cluster_markers.tsv"))

```

Plot them:

```{r seurat_markers_heatmap, fig.width = 7, fig.height = 18, dependson = 'process1'}

seurat_markers <- readr::read_tsv(glue("{out}-cluster_markers.tsv"))

DT::datatable(seurat_markers %>%
                group_by(cluster) %>%
                top_n(100, avg_logFC), filter = "top")

top15 <- seurat_markers %>% group_by(cluster) %>% top_n(15, avg_logFC)

DoHeatmap(object = seurat, genes.use = top15$gene,
          slim.col.label = TRUE, remove.key = TRUE,
          col.low = "#2166AC",
          col.mid = "#E5E0DC",
          col.high = "#B2182B")

rownames(seurat_markers) <- NULL

```

# Session info

```{r sinfo, cache = FALSE}

sessionInfo()

```