---
title: "Anderson et al - process Foxp1 data"
author: "Selin Jessa"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: show
    css: ../../../sjessa/from_hydra/misc/rmd_custom.css
    theme: flatly
    toc: yes
    toc_depth: 4
    number_sections: true
    df_print: paged
---


```{r config, warning = FALSE}

# **MODIFY THIS CHUNK**

project_id <- "anderson_data" # determines the name of the cache folder
doc_id     <- "01" # determines name of the subfolder of `figures` where pngs/pdfs are saved
out        <- paste0("../processed_data/", doc_id)
figout     <- paste0("figures/", doc_id, "/")
cache      <- paste0("~/tmp/", project_id, "/", doc_id, "/")

```

```{r setup, include = FALSE}

# NO NEED TO MODIFY THIS CHUNK

knitr::knit_hooks$set( echo_label = function(before, options, envir) {
  if ( before ) {
    # Do nothing
  } else sprintf('<br><span style="color:#1abc9c">~[output @ *%s*]~</span>',
                 paste0(options$fig.path, "/", options$label, "...") )
})

knitr::opts_chunk$set(message = TRUE,
                      warning = FALSE,
                      error = FALSE,
                      cache = TRUE,
                      cache.path = cache,
                      fig.path = figout,
                      fig.keep = "all",
                      dev = c("png", "pdf"),
                      cache.lazy = FALSE)

grDevices::pdf.options(useDingbats = FALSE)

options(knitr.table.format = "html") 
knitr::opts_knit$set(width = 1600)

```

***

# Overview

Load Drop-seq expression matrix obtained from GEO, and create a Seurat object in order
to run basic analyses (normalization, dimensionality reduction, clustering, derivation of cluster
markers, etc).

# Libraries

```{r libs, warning = FALSE, message = FALSE, echo = TRUE, cache = FALSE}

library(tidyr)
library(dplyr)
library(readr)
library(magrittr)
library(glue)
library(purrr)
library(ggplot2)
library(cytobox)
library(cowplot)
library(Seurat)

source("../../../sjessa/from_hydra/HGG-G34/samples/functions.R")
source("../../../sjessa/from_hydra/misc/sjlib.R")
ggplot2::theme_set(cytobox::theme_min(border_colour = "black", border_size = 0.5))

```

# Read data

```{r read_data}

# Read in the data into sparse matrix format
mat <- data.table::fread("../data/GSE125290_AAX_ALL_CLUST_RAW_COUNTS.txt.gz", data.table = FALSE)
rownames(mat) <- mat$CellBarcode

# Extract sample from cell ID (columns are cells with naming scheme {sample}_{cell barcode})
metadata <- mat[, 1:9]
mat <- mat[, -c(1:9)]
mat <- as.matrix(mat)

# Only keep cells in the CTL genotype
mat <- mat[metadata$Genotype == "CTL", ]
metadata <- metadata[metadata$Genotype == "CTL", ]

# Get cluster labels
cluster_lab <- readxl::read_xlsx("../data/Table_S2_cluster_labels_and_counts.xlsx") %>% 
  dplyr::select(Cluster_num = Cluster, Label = `Cell-type Annotation`) %>% 
  dplyr::mutate(Cluster = paste0(Cluster_num, "-", Label)) %>% 
  select(-Label)
metadata <- left_join(metadata %>% dplyr::rename(Cluster_num = Cluster), cluster_lab, by = "Cluster_num")
rownames(metadata) <- metadata$CellBarcode

# Sanity check
dim(mat)
dim(metadata)

```

# QC

## Create the Seurat object

```{r seurat}

seurat <- CreateSeuratObject(raw.data = t(mat),
                             min.cells = 3,
                             min.genes = 0,
                             is.expr = 0,
                             project = "Anderson_Foxp1_striatum",
                             names.field = 1,
                             names.delim = "_",
                             meta.data = metadata)

# Add mitochondrial content to the metadata
mito_genes <- grep("^mt-", rownames(seurat@data), value = TRUE)
percent_mito <- Matrix::colSums(seurat@data[mito_genes, ])/Matrix::colSums(seurat@data) * 100
seurat <- AddMetaData(seurat, percent_mito, "percent_mito")

# Add cell cycle score
seurat <- addCellCycle(seurat, species = "m_musculus")

```

## QC metrics

```{r qc_before_filt, fig.width = 9, fig.height = 5}

ngene_lim <- 10000
numi_lim  <- 25000
mito_lim  <- 10

plot_grid(VlnPlot(seurat, c("nGenes"),   
                  do.return = TRUE, point.size.use = -1, x.lab.rot = TRUE) + ylim(NA, ngene_lim),
          VlnPlot(seurat, c("nUMI"),          
                  do.return = TRUE, point.size.use = -1, x.lab.rot = TRUE) + ylim(NA, numi_lim),
          VlnPlot(seurat, c("percent_mito"),
                  do.return = TRUE, point.size.use = -1, x.lab.rot = TRUE) + ylim(NA, mito_lim),
          ncol = 3)

```

# Dimensionality reduction and clustering

```{r process1, results = "hide", dependson = 'thresholds'}

(process_params <- data.frame(
  n_pcs = 10,
  resolution = 0.6,
  seed = 100
))

seurat <- seurat %>%
  NormalizeData(normalization.method = "LogNormalize", scale.factor = 10000) %>%
  FindVariableGenes(mean.function = ExpMean, dispersion.function = LogVMR,
                    x.low.cutoff = 0.0125, x.high.cutoff = 3, y.cutoff = 0.5) %>%
  ScaleData(vars.to.regress = c("nUMI", "percent_mito")) %>%
  RunPCA(pc.genes = .@var.genes, pcs.compute = 100, do.print = FALSE) %>%
  RunTSNE(dims.use = 1:process_params$n_pcs, do.fast = TRUE, seed.use = process_params$seed) %>%
  FindClusters(dims.use = 1:process_params$n_pcs, resolution = process_params$resolution, random.seed = process_params$seed)

seurat <- set_qual_pal(seurat)
seurat <- SetAllIdent(seurat, "Cluster")

# Add in the UMAP coordinates from the paper
seurat@dr$umap <- seurat@dr$tsne
seurat@dr$umap@cell.embeddings[, 1] <- mat$UMAP_1
seurat@dr$umap@cell.embeddings[, 2] <- mat$UMAP_2

names(seurat@dr$umap@cell.embeddings) <- c("UMAP1", "UMAP2")

saveRDS(seurat, glue("{out}-seurat.Rds"))
write_tsv(process_params, glue("{out}-process_params.tsv"))

```

## Visualization and QC

How many cells per cluster?

```{r n_cells, dependson = 'process1'}

table(seurat@ident)

```

```{r plot_tsne_qc1, fig.width = 11, fig.height = 9, dependson = 'process1'}

plot_grid(cytobox::pca(seurat, title = "PCA", legend = TRUE),
          tsne(seurat, title = "tSNE", legend = TRUE),
          tsne(seurat, title = "nGenes", colour_by = "nGenes", colours = rdbu, colour_by_type = "continuous", label = FALSE),
          tsne(seurat, title = "nUMI", colour_by = "nUMI", colours = rdbu, colour_by_type = "continuous", label = FALSE),
          tsne(seurat, title = "percent_mito", colour_by = "percent_mito", colours = rdbu, colour_by_type = "continuous", label = FALSE),
          tsne(seurat, title = "G1_S_score", colour_by = "G1_S_score", colours = ylrd, colour_by_type = "continuous", label = FALSE),
          tsne(seurat, title = "G2_M_score", colour_by = "G2_M_score", colours = ylrd, colour_by_type = "continuous", label = FALSE))

```


<!-- ## PCA -->

<!-- ```{r viz_pca, fig.width = 12, fig.height = 4, dependson = 'process1'} -->

<!-- plot_grid(cytobox::pca(seurat, dim1 = 1, dim2 = 2), -->
<!--           cytobox::pca(seurat, dim1 = 1, dim2 = 3), -->
<!--           cytobox::pca(seurat, dim1 = 1, dim2 = 4), ncol = 3) -->

<!-- VizPCA(seurat, 1:4, nCol = 4, font.size = 0.9) -->

<!-- getVarianceExplained(seurat) -->

<!-- ``` -->

<!-- # Markers -->

<!-- ```{r seurat_markers, dependson = 'process1', eval = TRUE} -->

<!-- seurat_markers <- FindAllMarkers(seurat) -->

<!-- write_tsv(seurat_markers, glue("{out}-cluster_markers.tsv")) -->

<!-- ``` -->

<!-- Plot them: -->

<!-- ```{r seurat_markers_heatmap, fig.width = 7, fig.height = 18, dependson = 'process1'} -->

<!-- seurat_markers <- readr::read_tsv(glue("{out}-cluster_markers.tsv")) -->

<!-- DT::datatable(seurat_markers %>% -->
<!--                 group_by(cluster) %>% -->
<!--                 top_n(100, avg_logFC), filter = "top") -->

<!-- top15 <- seurat_markers %>% group_by(cluster) %>% top_n(15, avg_logFC) -->

<!-- DoHeatmap(object = seurat, genes.use = top15$gene, -->
<!--           slim.col.label = TRUE, remove.key = TRUE, -->
<!--           col.low = "#2166AC", -->
<!--           col.mid = "#E5E0DC", -->
<!--           col.high = "#B2182B") -->

<!-- rownames(seurat_markers) <- NULL -->

<!-- ``` -->

# Session info

```{r sinfo, cache = FALSE}

sessionInfo()

```