---
title: "05 - Plot 4C contacts at GSX2 and PDGFRA"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    keep_md: true
    theme: flatly
    css: ../../include/style.css
    toc: yes
    toc_depth: 4
    number_sections: true
    df_print: paged
    includes:
      before_body: ../../include/header.html
      after_body:  ../../include/footer.html
---

<!-- FRONT MATTER, insert configuration info -->

```{r header, echo = FALSE, results = 'asis', warning = FALSE}

# Index of the document
# ...determines name of the subfolder of `outputs` and `figures`
doc_id <- "05"
subdir <- "bulk_transcriptome_epigenome"

suppressMessages(library(here))

# Knit child document with header
res <- knitr::knit_child(here("include", "header.Rmd"), envir = environment(), quiet = TRUE)
cat(res, sep = '\n')

```

<!-- END OF FRONT MATTER -->


# Overview

In this document, we load a 4C quantification of the contacts between parts of the genome from Hi-C 
data in tumors and normal brain, focusing on contacts anchored at PDGFRA or GSX2.

# Libraries

```{r libs, warning = FALSE, message = FALSE, echo = TRUE, cache = FALSE}

library(readxl)
library(ggplot2)
library(cowplot)
library(tidyr)
library(dplyr)

# Theme
ggplot2::theme_set(theme_min(base_size = 13))

# Custom
source(here::here(subdir, "analysis/functions.R"))

```

# Analysis

## Anchored at PDGFRA

Generate the plots for contacts anchored at the PDGFRA promoter:

### Human

```{r prep_human}

data <- read_excel(here(subdir, "data/PDGFRA_promoter_4C_human.xlsx"))

# Convert to long format so that we have three columns in the data frame:
# 1) Group (G34 or Other), 2) Average value, and 3) Standard deviation value
data_long <- data %>%
    unite(G34,   Average_G34_pGBMs,   stdeva_G34_pGBMs) %>%
    unite(Other, Average_Other_pGBMs, stdeva_other_pGBMs) %>%
    gather(Group, Value, G34, Other) %>%
    separate(Value, into = c("Average", "Stdev"), sep = "_")

# Convert the missing values around PDGFRA to 0
data_long$Average <- ifelse(data_long$Average == "NA", 0, data_long$Average)
data_long$Stdev   <- ifelse(is.na(data_long$Stdev),    0, data_long$Stdev)

# Make sure we have numeric columns for plotting
data_long$Average <- as.numeric(data_long$Average)
data_long$Stdev   <- as.numeric(data_long$Stdev)

```


```{r human_pdgfra_plot, fig.width = 7, fig.height = 3, echo_fig = TRUE}

# Colour palette
palette_human <- c("red", "navy")

# Write a plotting function
plot_4c <- function(df, palette) {
  
  df %>%
    rr_ggplot(aes(x = START, y = Average), plot_num = 1) +
    # Plot the stdev as an area "ribbon"
    geom_ribbon(aes(ymin = Average - Stdev, ymax = Average + Stdev, fill = Group), alpha = 0.3) +
    # Plot the solid line for the average
    geom_line(aes(colour = Group)) +
    # Set the colours
    scale_colour_manual(values = palette) +
    scale_fill_manual(values = palette) +
    # Add more breaks to the x-axis
    scale_x_continuous(breaks = seq(from = min(data$START), to = max(data$START), by = 20000)) +
    theme_cowplot() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = rel(0.7)))
  
}

plot_4c(data_long, palette_human)

```


### Mouse

```{r prep_mouse}

data <- read_excel(here(subdir, "data/PDGFRA_promoter_4C_mice.xlsx"))

# Convert to long format so that we have three columns in the data frame:
# 1) Group (G34 or Other), 2) Average value, and 3) Standard deviation value
data_long <- data %>%
  unite(GE,     Average_GE, STDEVA_GE) %>%
  unite(Cortex, Average_Cortex, STDEVA_Cortex) %>%
  unite(ESC,    Average_ESC, STDEVA_ESC) %>%
  gather(Group, Value, GE, Cortex, ESC) %>%
  separate(Value, into = c("Average", "Stdev"), sep = "_")

# Make sure we have numeric columns for plotting
data_long$Average <- as.numeric(data_long$Average)
data_long$Stdev   <- as.numeric(data_long$Stdev)

```


```{r mouse_pdgfra_plot, fig.width = 7, fig.height = 3, echo_fig = TRUE}

# Colour palette
palette_mouse <- c("blue", "orange", "gray50")

# Plot
plot_4c(data_long, palette_mouse)

```

## Anchored at GSX2

Generate the plots for contacts anchored at the GSX2 promoter:

### Human

```{r prep_human_gsx2}

data <- read_excel(here(subdir, "data/GSX2_promoter_4C_human.xlsx"))

# Convert to long format so that we have three columns in the data frame:
# 1) Group (G34 or Other), 2) Average value, and 3) Standard deviation value
data_long <- data %>%
    unite(G34,   Average_G34, STDEVA_G34) %>%
    unite(Other, Average_other_pGBMs, STDEVA_non_G34) %>%
    gather(Group, Value, G34, Other) %>%
    separate(Value, into = c("Average", "Stdev"), sep = "_")

# Make sure we have numeric columns for plotting
data_long$Average <- as.numeric(data_long$Average)
data_long$Stdev   <- as.numeric(data_long$Stdev)

```


```{r human_gsx2_plot, fig.width = 7, fig.height = 3, echo_fig = TRUE}

plot_4c(data_long, palette_human)

```


### Mouse

```{r prep_mouse_gsx2}

data <- read_excel(here(subdir, "data/GSX2_promoter_4C_mice.xlsx"))

# Convert to long format so that we have three columns in the data frame:
# 1) Group (G34 or Other), 2) Average value, and 3) Standard deviation value
data_long <- data %>%
  unite(GE,     Average_GE, STDEVA_GE) %>%
  unite(Cortex, Average_Cortex, STDEVA_Cortex) %>%
  unite(ESC,    Average_ESC, STDEVA_ESC) %>%
  gather(Group, Value, GE, Cortex, ESC) %>%
  separate(Value, into = c("Average", "Stdev"), sep = "_") %>% 
  filter(Group != "ESC")

# Make sure we have numeric columns for plotting
data_long$Average <- as.numeric(data_long$Average)
data_long$Stdev   <- as.numeric(data_long$Stdev)

```


```{r mouse_gsx2_plot, fig.width = 7, fig.height = 3, echo_fig = TRUE, dependson = 'prep_mouse_gsx2'}

plot_4c(data_long, palette_mouse)

```


<!-- END MATTER, insert reproducibility info -->

```{r footer, echo = FALSE, results = 'asis', warning = FALSE, cache = FALSE}

# Knit child document with header
res <- knitr::knit_child(here::here("include", "footer.Rmd"), envir = environment(), quiet = TRUE)
cat(res, sep = '\n')

```

<!-- END OF END MATTER -->
